\documentclass[12pt]{article}
    \usepackage[margin=1in]{geometry}
    \usepackage{mdframed}
    \usepackage{subcaption}
    \usepackage{amssymb}
    \usepackage{amsmath}
    \usepackage{mathtools}

    % Custom colors
    \usepackage{color}
    \definecolor{deepblue}{rgb}{0,0,0.5}
    \definecolor{deepred}{rgb}{0.6,0,0}
    \definecolor{deepgreen}{rgb}{0,0.5,0}
    % Default fixed font does not support bold face
    \DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{12} % for bold
    \DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{12}  % for normal
    \usepackage{listings}
    \usepackage{url}

    \usepackage{hyperref}

    \usepackage{booktabs}
    
    \usepackage{tikz}
    \usetikzlibrary{shapes,arrows,automata,positioning,cd}
    \tikzset{
      cfgedge/.style   = {black, ->, >=stealth},
      forward/.style = { blue, ->, >=angle 45},
      backward/.style = { red, densely dashed, ->, >=latex' },
      backwardleft/.style = { red, densely dashed, <-, >=latex' },
    }
    \usepackage{xcolor}
    
    \newcommand{\cfgarrow}{\mathbin{\tikz[baseline]\draw[cfgedge,yshift=0.6ex] (0,0) -- (.9em,0);}}
    \newcommand{\forwardarrow}{\mathbin{\tikz[baseline]\draw[forward,yshift=0.6ex] (0,0) -- (1em,0);}}
    \newcommand{\backwardarrow}{\mathbin{\tikz[baseline]\draw[backward,yshift=0.6ex] (0,0) -- (.95em,0);}}
  \newcommand{\dom}{\underline{\gg}}
    
    
    \begin{document}
    \lstset{
    language=C,
    basicstyle=\ttfamily\small,
    keywordstyle=\ttb\color{deepblue}\small,
    emph={foo,bar,assert,baz},          % Custom highlighting
    emphstyle=\ttb\color{deepred}\small,    % Custom highlighting style
    stringstyle=\color{deepgreen}\small,
    frame=tb,                         % Any extra options here
    numbers=left,
    stepnumber=1,
    showstringspaces=false            % 
    }
    
    \begin{center}
        \bigskip
        {\LARGE ECS 240 Programming Languages} \medskip
                
        {\Large Homework 2} \bigskip
    
    \end{center}

    \section*{About This Assignment}
    
    \begin{itemize}
      \item This assignment tests you on your understanding of points-to analysis.
      \item To complete the assignment (i)~modify \texttt{hw2.tex}, (ii)~create
      the corresponding PDF document (using pdflatex, for example), and
      (iii)~submit the pdf electronically via Gradescope by the due date; see
      \href{https://www.gradescope.com/get_started#student-submission}{this page}
      and
      \href{http://gradescope-static-assets.s3-us-west-2.amazonaws.com/help/submitting_hw_guide.pdf}{this
      document}. Your assignment will not be graded and you will receive no
      points if you do not follow these instructions. 

      \item \textbf{When submitting the assignment on Gradescope, please mark
      which page corresponds to each question on the assignment.}  Your
      assignment will not be graded and you will receive no points if you do not
      mark the pages.      

      \item The  \LaTeX\ source has \texttt{TODO}~comments to clearly indicate
      where changes need to be made. 
      \item The \verb=\vspace= commands can be safely commented out.
      \item This assignment can be worked on in a group of at most three. Enter
      the names and email addresses of the team members in the space provided
      below.
    \end{itemize}

    \begin{mdframed}
      Team members:
      \begin{itemize}
        %TODO
        \item Name, email address: Soham Kolhatkar, sakolhatkar@ucdavis.edu
        \item Name, email address: Divyansh Rajesh Jain, drajeshjain@ucdavis.edu
        %\item Name, email address % Comment out this line if working in pairs.
      \end{itemize}
    \end{mdframed}
    
    \newpage
    \begin{enumerate}
        \item (5 points) Consider the following C function:
        \begin{lstlisting}
          void bar(int *p, int *q) {
            if (*p != *q) return;
            (*q)++;
            (*p)++;
            (*q)++;
            // P1
          }
        \end{lstlisting}
        A program analyzer $A$ reports the following expression as an invariant
        at program point \lstinline$P1$: \lstinline$*p < *q$.

        Based solely on this report, state whether the following statement is
        \textbf{True} or \textbf{False}: 

        \emph{The program analyzer $A$ is sound.}

        If \textbf{True}, explain your reasoning. If \textbf{False}, provide an
        input for \lstinline$bar$ for which the expression evaluates to false.
         \begin{mdframed}
       
        %% TODO Comment out the correct answer.
        %\textbf{True} 
        \textbf{False}

        An input for which the invariant \lstinline$*p < *q$ is false is:
        
        int* p = \&a; 
        
        int* q = \&a;

        Where $p$ and $q$ point to the same variable $a$. Then by statment P1, p will equal q.
        \end{mdframed}

        \item (5 points) Write down the set of statements $S$ such that performing precise flow-insensitive points-to
        analysis results in the following realizable pairs:\\
        $\{\langle a, r\rangle,\langle p, b\rangle,  \langle a, b\rangle, \langle p, r\rangle\}$.
        The size of $S$ should be \emph{at most 3}.
        \begin{mdframed}
          %% TODO At most three statements in S.
          $S = \{ a = \&b  ,  a = \&r  , p = a  \}$
          \end{mdframed}


        \item (5 points) Write down a program $P$ such that the output of a precise
        flow-sensitive points-to analysis at the end of the program $P$ is more
        precise compared to that of a precise flow-insensitive points-to
        analysis for the same program; that is, there are fewer points-to relations 
        inferred by flow-sensitive points-to analysis at the end of $P$ compared 
        to those inferred by precise flow-insensitive points-to analysis.
        
        Apart from the program $P$, your answer should also list the points-to 
        relations inferred by flow-sensitive and flow-insensitive analysis.
        
     \vspace{12em}

        \begin{mdframed}
         
          %% TODO
           \textbf{Program P:}
            \begin{lstlisting}
            #include <stdio.h>
            
            int main() {
                int a = 5;
                int b = 6;
                int c = 7;
            
                int* p = &a;
                p = &b;
                p = &c;
            
                printf("%d\n", *p); //  p only points to c
                return 0;
            }
            \end{lstlisting}
          Points-to relations at end of P using flow-sensitive analysis:

        \[
        \{ \langle p, c \rangle \}
        \]

          Points-to relations using flow-insensitive analysis

         \[
        \{ \langle p, a \rangle, \langle p, b \rangle, \langle p, c \rangle \}
        \]

        \end{mdframed}

      \item (5 points) A statement involving multiple dereferences can be broken
      into multiple statements each with a single deference involving temporary
      variables. Given a program $P$, let $OneStar(P)$ denote such a rewrite.
      For example, if $P \coloneqq$ \lstinline$y = &z; x = ***y$, then
      $OneStar(P) \coloneqq $ \lstinline$y = &z; t2 = *y; t1 = *t2; x = *t1;$.
      This rewrite does not reduce the precision of flow-sensitive points-to
      analysis. However, it could lead to imprecision for flow-insensitive
      points-to analysis; that is, performing precise flow-insensitive analysis
      on $P$ could produce fewer points-to relations compared to performing
      precise flow-insensitive analysis on $OneStar(P)$ even when we just
      consider the points-to relations among the variables of $P$.

      Give an example of a program $P$ such that precise flow-insensitive
      points-to analysis of $P$ is more precise compared to that for
      $OneStar(P)$ when considering \emph{only} the variables in $P$. Apart from
      $P$, your answer should also list $OneStar(P)$ and the points-to relations
      corresponding to $P$ and $OneStar(P)$.
      
      \begin{mdframed}
        \vspace{2em}
        %% TODO
        %% Program P
        %% Points-to relations for P using flow-insensitive analysis
        %% Program OneStar(P)
        %% Points-to relations for OneStar(P) using flow-insensitive analysis
      \end{mdframed}

      \item (5 points) Recall that a \emph{program with well-defined types}
      contains only variables that are scalars with well-defined types (e.g.
      \lstinline$int$, \lstinline$int*$, \lstinline$int**$, etc.). Furthermore,
      a variable can only point to variables of compatible type. For example, a
      variable of type \lstinline$int**$ can only point to a variable of type
      \lstinline$int*$, and a variable of type \lstinline$int*$ can only point
      to a variable of type \lstinline$int$.
      
      For a given program \textbf{P}, the following \emph{realizability graph}
      $G_R$ is constructed by computing precise flow-insensitive points-to
      analysis for \textbf{P}:
      \vspace{-2ex}
      \begin{center}
      \begin{tikzpicture}[auto,
        node distance=1.4cm, 
        ] 
        \tikzstyle{every node} = [circle, draw]
        \node (p1) {$p_1$};
        \node [right of=p1] (p2) {$p_2$};
        \node [below of=p1] (x) {$x$};
        \node [below of=p2] (y) {$y$};
        \node [below of=x] (a) {$a$};
        \node [below of=y] (b) {$b$};
        
        \path (p1) edge[cfgedge] (x); 
        \path (p1) edge[cfgedge] (y); 
        \path (p2) edge [cfgedge] (x);
        \path (x) edge[cfgedge] (y); 
        \path (x) edge [cfgedge] (a);
        \path (y) edge [cfgedge] (b);    
        
        \end{tikzpicture}
      \end{center}
      \vspace{-1ex}
        Based on $G_R$, mark (using $\checkmark$) \textbf{EXACTLY ONE} statement
        from below that is true for program~\textbf{P}, and \textbf{justify your
        answer}:
        \vspace{-2ex}
        \renewcommand{\arraystretch}{1.5}
        \begin{center}
        \begin{tabular}{|l|l|}
          \hline
          The program \textbf{P} might be a program with well-defined types. & \hspace{3em} \\ \hline
          The program \textbf{P} is definitely NOT a program with well-defined types. & $\checkmark$
          \\ \hline
        \end{tabular}
      \end{center}
      \vspace{-2ex}
       \begin{mdframed}
       
        As stated in the question, well-defined types are those that \textbf{only} point to compatible types ( int* points to int, etc.). In the realizability graph $G_R$ we know from class that $p_1$ and $p_2$ are equivalent types, that is $Type(p_1)$ = $Type(p_2)$. We also know that for $G_R$ to be defined with well-defined types, all nodes in any Layer $i$ must point to Layer $i-1$ \textbf{only}.

        However we can see that the node $x$ has a points-to relation with an adjacent node in the same layer. This means $G_R$ does not contain well-defined types. We also know that the $G_R$ is an over approximation of the program $P$. Therfore, we can say with certainty that the program $P$ is definitely NOT a program with well-defined types.
        
       \end{mdframed}
    

    \end{enumerate}
    
\end{document}